---
title: "Graining level used in studies"
author: "Mariia Bilous"
date: "2023-03-21"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
```

Read table with collected graining levels across studies
```{r}
gamma.df <- read.csv("graining_level.csv")
```

Filter out irrelevant methods (i.e., hdWGCNA, which use neighborhood approach instead of metacells) and group rare tools 
```{r}
tools <- unique(gamma.df$metacell_tool)
tools

keep.tools <- c("Other", "SEACell", "SuperCell", "MetaCell")   
colors <- c("#f2e9e4", "#9a8c98", "#4a4e69", "#22223b")
names(colors) <- c(keep.tools)

gamma.df <- gamma.df %>%
  filter(metacell_tool != "hdWGCNA") # not a metacell tool (neighborhoods)

gamma.df$metacell_tool_exact <- gamma.df$metacell_tool

gamma.df$metacell_tool[!(gamma.df$metacell_tool %in% keep.tools)] <- "Other"
gamma.df$metacell_tool <- factor(gamma.df$metacell_tool, rev(keep.tools))
```

Overall gamma distribution and statistics
```{r}
hist(log10(gamma.df$graining_level))

summary(gamma.df$graining_level[gamma.df$metacell_tool != 'MetaCell'])
summary(gamma.df$graining_level[gamma.df$metacell_tool == 'MetaCell'])
summary(gamma.df$graining_level)
```

Plot gamma distribution colored by metacell tool 

```{r}
gamma.df %>%
  ggplot(aes(x = graining_level, fill = metacell_tool)) + 
  geom_histogram(alpha = 1, color = "white") +
  scale_fill_manual(values = colors) +
  scale_x_log10() +
  labs(x = "Graining level", y = "Counts") +
  theme_classic() +  theme(asp = 0.70) 

ggsave(file = "./plots/gamma_distr.pdf", width = 6, height = 6)

```

Plot correlation between a size of dataset and applied graining level

```{r}
p <- gamma.df %>%
  ggplot(aes(x = graining_level, y = N_cells )) + 
  geom_point(aes(color = metacell_tool), alpha = 0.9) +
  scale_fill_manual(values = colors)+
  scale_color_manual(values = colors)+
  theme(legend.position = "bottom") +
  scale_y_log10() + scale_x_log10() +
  labs(x = "Graining level", y = "Dataset size") +
  theme_classic()+
  theme(asp = 0.7) +
  geom_smooth(method='lm', color = NA, se = FALSE)

p1 <- p + ggpubr::stat_cor(method="pearson", size = 3)
p1
ggsave("./plots/gamma_distr_scatter.pdf", width = 6, height = 6)

p2 <- p + ggpubr::stat_cor(aes(color = metacell_tool), method="pearson", p.accuracy = 0.001, r.accuracy = 0.01, size = 3)
p2
ggsave("./plots/gamma_distr_scatter_per_method_correlation.pdf", width = 6, height = 6)
```


## Metacell size distribution scheme

```{r mc-size-distribution}
set.seed(12345)
s2 <- round(rnorm(100, 20, 5)) 
s1 <- round(rnorm(100, 20, 10))

s2[s2<1] <- 1

df <- data.frame(size = c(s1, s2), group = rep(c("s1", "s2"), c(length(s1), length(s2))))

df %>% ggplot(aes(y = size, x = factor(group))) + geom_violin(fill = "gray", trim = F, adjust = 1) + theme_void()
ggsave("./plots/MC_size_distribution_scheme.pdf")
```



